Binary to continuous methodology draft

### 2024-07-26 notes/alloscore-application.Rnw
more cleanup

All forecast data used in this evaluation are available through the COVID-19 Forecast Hub (cite repository). Code for our analyses is available on GitHub (cite repository).

>>>>

All forecast data used in this evaluation are available through the COVID-19 Forecast Hub \citep{cramer_reichlabcovid19-forecast-hub_2021}. An R package implementing the allocation score is available at \url{https://github.com/aaronger/alloscore} and all code for the analyses presented in this manuscript is available at \url{https://github.com/aaronger/utility-eval-papers}.

### 2024-07-26 notes/alloscore-application.tex
more cleanup

\subsection{Data}
<<get-data-dates>>=
metadata <- tar_meta()
@

\subsubsection{Hospitalization data}
Starting in the summer of 2020, the US Health and Human Services (HHS) began reporting counts of daily new admissions to hospitals for individuals with COVID-19 \citep{healthdatagov_covid-19_nodate}.
These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data \cite{wang-covidhubutils}.
The hospitalization data were downloaded for analysis on \Sexpr{format(as.Date(metadata$time[which(metadata$name == "truth_data")]), "%B %d, %Y")}.

@@

<<ias-comparison>>=
tar_load(alloscore_2021.12.20)
slim_alloscore_2021.12.20 <- alloscore::slim(alloscore_2021.12.20, id_cols = c("model", "reference_date"))

## for visual context
## curve(dnorm(x, mean=15000, sd=3000), from=0, to=30000)
ks <- seq(5000, 25000, by=200)
kprobs <- data.frame(K = ks,
                     weight = dnorm(ks, mean=15000, sd=3000)/sum(dnorm(ks, mean=15000, sd=3000)))

ias <- slim_alloscore_2021.12.20 |>
  group_by(model, K) |>
  summarize(`Allocation score` = sum(components)) |>
  left_join(kprobs) |>
  group_by(model) |>
  summarize(IAS_wt = sum(`Allocation score`*weight, na.rm=TRUE),
            IAS = mean(`Allocation score`)) |>
  arrange(IAS)

as <- slim_alloscore_2021.12.20 |>
  filter(K == 15000) |>
  group_by(model) |>
  summarize(AS = sum(components)) |>
  arrange(AS)
@

<<multi-k-scores, results='asis'>>=
wis_vals <- all_scores |>
  filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
  select(model,
# can uncomment the two lines below to check they match
#         AS = alloscore,
#         `AS rank` = allo_rank,
         MWIS = mwis)

left_join(ias, as) |>
  left_join(wis_vals, by="model") |>
  rename(`IAS centered at 15k` = IAS_wt,
         `IAS uniform` = IAS) |>
  select(model, AS, MWIS, `IAS centered at 15k`, `IAS uniform`) |>
  arrange(AS) |>
  xtable::xtable(caption = "Comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.",
                 digits = 0,
                 label = "tab:multi-k-scores") |>
  print(include.rownames = FALSE, )
@

<<metrics-rankings, results = "asis", eval=FALSE>>=
## NOTE: this content now shown in multi-k-scores chunk
## for highlighting cells, if we want that https://stackoverflow.com/questions/37448348/color-code-cells-of-xtable
all_scores |>
  filter(target_end_date == "2022-01-03", !is.na(mwis)) |>
  select(model,
         AS = alloscore,
         `AS rank` = allo_rank,
         MWIS = mwis,
         `MWIS rank` = mwis_rank) |>
  arrange(`AS rank`) |>
  xtable::xtable(label = "tab:one-week-results",
                 caption = "Allocation scores (AS) and mean weighted interval scores (MWIS) for each model' as forecasted's forecasts for ") |>
  print(include.rownames = FALSE)
@

@@

<<detailedASinfo>>=
source("R/run-alloscore.R")
as_Buck <- run_alloscore(
  forecast_data = forecast_data,
  truth_data = truth_data,
  one_reference_date = "2021-12-20",
  one_model = "JHUAPL-Bucky",
  one_K = 15000,
  slim = FALSE)

as_CU <- run_alloscore(
  forecast_data = forecast_data,
  truth_data = truth_data,
  one_reference_date = "2021-12-20",
  one_model = "CU-select",
  one_K = 15000,
  slim = FALSE)

Buck_quantile <- round(1 - as_Buck$lam,2)*100
CU_quantile <- round(1 - as_CU$lam,2)*100

FL_allocs <- Kat15k_alloscores |>
   dplyr::filter(
     model %in% c("JHUAPL-Bucky", "CU-select"),
     reference_date == "2021-12-20",
     abbreviation == "FL") |>
  dplyr::select(model, x) |>
  tibble::deframe()
FL_diff <- round(FL_allocs["CU-select"] - FL_allocs["JHUAPL-Bucky"])
@

For JHUAPL-Bucky, the allocations were assigned at the \Sexpr{Buck_quantile}th percentile of the predictive distribution and for CU-select they were at the \Sexpr{CU_quantile}th percentile.
This reflects that in general the JHUAPL-Bucky forecasts assigned more probability to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
In California, New York and Texas, the eventually observed number of hospital admissions was closer to the predicted median from JHUAPL-Bucky than the median from CU-select, and was contained in JHUAPL-Bucky's 80\% PI but not in the 80\% PI from CU-select. However for each of these three states, the resource allocation was lower for JHUAPL-Bucky than for CU-select and therefore resulted in more hypothetical unmet need.
In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was \Sexpr{FL_diff} units lower than the allocation derived from the CU-select model.

<<thermometer-plot, fig.height=3.5, fig.width=7.5, fig.cap="Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts. For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts. The side-by-side shaded regions show the median (solid horizontal line) and 50%, 80% and 95% prediction intervals for the two selected models. The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data). The vertical bars with red and yellow shading show the allocations: the red bar goes from zero to the level of the observed number of hospitalizations while the yellow bar shows how many resources were suggested by the model to be allocated to that location. The allocations (yellow bar) may be more or less than the observed number, therefore exposed red bar indicates unmet need, while the yellow bar extending above the observed number of hospitalizations means that excess resources were suggested for that given location.">>=
source("R/plot_functions.R")
plot_hosp(models = c("JHUAPL-Bucky", "CU-select"),
          start_date = "2021-12-01",
          stop_date = "2022-01-25",
          f_date = "2021-12-20",
          te_date = "2022-01-03",
          ## highest locations in terms of hospitalizations
          locations = c("FL", "NY", "TX", "CA"),
          f_width1 = 4, f_alpha = .4,
          allocations = TRUE,
          forecasts_hosp = forecast_data %>%  filter(reference_date == "2021-12-20"),
          alloscores = slim_alloscore_2021.12.20,
          truth = truth_data,
          f_colors = c("#4A708B", "#CD2626"),
          one_K = 15000) + labs(x = NULL)
@

@@

<<multi-loc-ranks, fig.cap="Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select). Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model. The number of available resources was fixed at 15,000 and forecasts from each model were used to determine an optimal allocation strategy before the resource need was known. For most locations the resource need exceeded the resources allocated, indicated by the `observed resource need' bar being larger than the `allocation' bar. Panel B shows the breakdown of the weighted interval score (WIS) into components of underprediction, overprediction and dispersion. Larger values of WIS indicate more error, and the full WIS score for each location can be decomposed into the three components shown here.">>=
## date points to target_end_date of 2022-01-10 to highlight
ref_date_for_plot <- "2021-12-20"
models_to_plot <- c("JHUAPL-Bucky", "CU-select")

multiloc_dat <- Kat15k_alloscores |>
  filter(reference_date == ref_date_for_plot,
         ## the above date points to target_end_date of 2022-01-03 where things were really wrong
         model %in% models_to_plot) |>
  rename(allocation = x, `observed resource need` = y, state = abbreviation) |>
  select(model, reference_date, state, allocation, `observed resource need`) |>
  pivot_longer(cols = c("allocation", "observed resource need"))

## plot all locations observations and allocations

## line plot below here
# multiloc_plot <- ggplot(multiloc_dat, aes(x=name, y=value, group=state)) +
#   geom_line() +
#   geom_line(data=dplyr::filter(multiloc_dat, state=="FL"), color="orange") +
#   ylab("Daily hospital admissions") +
#   scale_x_discrete(name=NULL, limits=rev, expand = c(.1,.1)) +
#   facet_grid(.~model)

ordered_levels <- truth_data |>
  filter(target_end_date == as.Date(ref_date_for_plot)+14, location != "US") |>
  select(location, abbreviation, value) |>
  arrange(-value)

alloc_plot <- multiloc_dat |>
  mutate(state = factor(state, levels = ordered_levels$abbreviation),
         name = factor(name, levels = c("observed resource need", "allocation"))) |>
  arrange(name) |>
  ggplot(aes(x=state, y=value, fill=name)) +
  scale_fill_manual(name=NULL, values = c("#CD2626", "goldenrod")) +
  theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.9,.95), legend.justification = c(1,1)) +
  ylab("Resource need and allocation") +
  xlab(NULL) +
  geom_col(position="identity", alpha=.7) +
  facet_grid(model~.)

wis <- covidHubUtils::score_forecasts(
  forecast_data |> filter(reference_date == "2021-12-20", str_detect(model,"Buck|CU-select")),
  truth_data,
  use_median_as_point = TRUE,
  metrics = "wis_components",
  return_format="long") |>
  filter(location != "US", score_name != "true_value") |>
  left_join(select(ordered_levels, location, abbreviation)) |>
  mutate(abbreviation = factor(abbreviation, levels=ordered_levels$abbreviation),
         `WIS component` = factor(score_name, levels = c("overprediction", "dispersion", "underprediction")))

## palette from
## colorspace::divergingx_hcl(palette = "Blue-Yellow 3", n=3)

wis_plot <- ggplot(data = wis, aes(x=abbreviation)) +
  geom_col(aes(y=score_value, fill=`WIS component`)) +
  scale_fill_manual(values = c("#9FA2FF", "gray77", "#BAAE00")) +
  facet_grid(model~.) +
  theme(axis.text.x=element_text(angle=90,vjust=.5), legend.position = c(.9,.95), legend.justification = c(1,1)) +
  xlab(NULL) +
  ylab("WIS")

cowplot::plot_grid(alloc_plot, wis_plot,
                   nrow=2,
                   align="v",
                   rel_heights = c(1, 1),
                   labels = "AUTO")

@

\subsubsection{Forecast scores showed differences over time}
<<grob-hosps-over-time>>=
dat_to_plot <- truth_data |>
  filter(location == "US",
         target_end_date %in% score_data_summaries$target_end_date)
p1 <- ggplot(dat_to_plot,
             aes(x=target_end_date, y=value)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept=15000, linetype=2) +
  ylab("hospital admissions") +
  xlab(NULL) +
  ylim(c(0, NA))
@

<<metrics-over-time, fig.height=9, fig.width=7, fig.cap="Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the assumed resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The x-axis corresponds to the date for which the prediction was made. AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
p2 <- ggplot(alloscore_data_summaries,
             aes(x=target_end_date, y=alloscore, color=model, shape=model)) +
  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
  ylab("Allocation score") +
  geom_point() + geom_line()+
  theme(legend.pos="none") +
  xlab(NULL)

p4 <- ggplot(score_data_summaries,
             aes(x=target_end_date, y=mwis, color=model,shape=model)) +
  geom_point() + geom_line() +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4)) +
  ylab("MWIS")+
  theme(legend.pos="bottom")+
  xlab(NULL)

cowplot::plot_grid(p1, p2, p4,
                   nrow=3,
                   align="v",
                   rel_heights = c(1, 1, 1.5),
                   labels = "AUTO")
@

@@

<<metrics-correlation, fig.width=10, fig.height=7.5, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week. Each facet of the plot corresponds to one model. Within each facet, each point corresponds to a week. The x- and y-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week. Points corresponding to earlier dates have darker shading. The size of the point corresponds to the observed value on the date for which the prediction was made. Models show different degrees of association between the two metrics.">>=
all_scores |>
  filter(as.Date(reference_date.x) < as.Date("2022-02-01") ) |>
  ggplot(aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
  geom_point(alpha=.7) +
  scale_x_continuous("MWIS standardized rank", limits=c(0,1)) +
  scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
  facet_wrap(.~model) +
  geom_abline(slope=1, intercept=0, linetype=2, color="gray")

@

@@

<<calc-tot-y>>=
tot_y <- slim_alloscore_2021.12.20 |>
  filter(model == "COVIDhub-ensemble", K==200) |>
  pull(y) |>
  sum()
@

AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 (Figure \ref{fig:multi-k}A).
These calculations highlight that AS was highest at the $K$ nearest to the observed value of \Sexpr{format(tot_y, big.mark = ",")}.
Rankings of AS from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.

@@

<<multi-k, fig.height=10, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20. Panel A shows, for each model, the AS for values of K between 200 and 60,000 at increments of 200. The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations. Panel B shows two possible weighting functions for the Integrated Allocation Score (IAS). The first (dark blue circles) computes weights proportional to a normal distribution centered at 15,000 (solid vertical gray line) with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000. The second (light orange triangles) uses a uniform weight for all possible values of K. Note that the AS used in earlier sections of the application uses the single fixed value of K=15,000.  Panel C shows how either method of computing the IAS (y-axis) is correlated with AS at K=15,000 (x-axis). Every point represents the AS and IAS for one model. The IAS centered at 15,000 are more closely correlated with the AS values.">>=

ias_to_plot <- left_join(ias, as) |>
  pivot_longer(cols = c(IAS, IAS_wt), names_to = "IAS_type", values_to = "IAS") |>
  mutate(IAS_type = ifelse(IAS_type=="IAS", "uniform", "centered at 15,000"))

p1 <- ggplot(ias_to_plot, aes(x=AS, y=IAS, color=IAS_type, shape=IAS_type)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, linetype=2) +
  scale_color_manual(name = "Integrated Allocation Score", values = c("#386cb0", "#fdc086")) +
  scale_shape_manual(name = "Integrated Allocation Score", values = c(1, 2)) +
  geom_abline(slope=1, intercept=0, color="gray") +
  theme(legend.position = c(0.05,.9), legend.justification = c(0,1),
        legend.background=element_rect(fill = alpha("white", 0.2)))

p2 <- slim_alloscore_2021.12.20 |>
  group_by(model, K) |>
  summarize(`Allocation score` = sum(components)) |>
  left_join(kprobs) |>
  ggplot(aes(x=K)) +
  geom_point(aes(y=weight), color = "#386cb0") +
  geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="#fdc086", shape = 2) +
  geom_vline(xintercept = 15000, color="gray")

p3 <- slim_alloscore_2021.12.20 |>
  group_by(model, K) |>
  summarize(`Allocation score` = sum(components)) |>
  mutate(model = factor(model, levels = rev(ias$model))) |>
  ggplot(aes(x=K, y=`Allocation score`, color=model, linetype=model)) +
  geom_line() +
  theme(legend.position = c(0.01,.99), legend.justification = c(0,1),
        legend.background=element_rect(fill = alpha("white", 0.2)),
        legend.title = element_blank(), legend.text = element_text(size=7))

cowplot::plot_grid(p3, p2, p1,
                   nrow=3,
                   align = "h",
                   labels = "AUTO")
@

>>>>

\subsection{Data}

\subsubsection{Hospitalization data}
Starting in the summer of 2020, the US Health and Human Services (HHS) began reporting counts of daily new admissions to hospitals for individuals with COVID-19 \citep{healthdatagov_covid-19_nodate}.
These daily counts were available for the US as a whole, and all states and several additional jurisdictions such as Puerto Rico and Washington DC.
The data were updated daily and were available for download by the public through the HHS HealthData.gov website.
For this analysis, we downloaded the hospitalizations data through the covidHubUtils R package, which connects users to the most recent version of the data \cite{wang-covidhubutils}.
The hospitalization data were downloaded for analysis on December 03, 2023.

@@

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Thu Dec 21 23:53:38 2023
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
  \hline
CU-select & 669 & 133 & 774 & 326 \\ 
  COVIDhub-ensemble & 873 & 159 & 1067 & 438 \\ 
  USC-SI\_kJalpha & 995 & 91 & 1216 & 1097 \\ 
  JHUAPL-Gecko & 1034 & 164 & 1141 & 418 \\ 
  MUNI-ARIMA & 1084 & 169 & 1248 & 440 \\ 
  COVIDhub-trained\_ensemble & 1089 & 169 & 1271 & 823 \\ 
  COVIDhub-baseline & 1175 & 170 & 1317 & 535 \\ 
  JHUAPL-Bucky & 1358 & 102 & 1566 & 1214 \\ 
  JHUAPL-SLPHospEns & 1540 & 129 & 1604 & 1102 \\ 
  UVA-Ensemble & 2469 & 213 & 2635 & 2494 \\ 
   \hline
\end{tabular}
\caption{Comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.} 
\label{tab:multi-k-scores}
\end{table}

@@

For JHUAPL-Bucky, the allocations were assigned at the 28th percentile of the predictive distribution and for CU-select they were at the 89th percentile.
This reflects that in general the JHUAPL-Bucky forecasts assigned more probability to higher values and thus a low quantile value across all locations reached the allocation limit of 15,000.
In California, New York and Texas, the eventually observed number of hospital admissions was closer to the predicted median from JHUAPL-Bucky than the median from CU-select, and was contained in JHUAPL-Bucky's 80\% PI but not in the 80\% PI from CU-select. However for each of these three states, the resource allocation was lower for JHUAPL-Bucky than for CU-select and therefore resulted in more hypothetical unmet need.
In Florida, which was the state that saw the largest number of reported hospitalizations on this target date, the forecast from JHUAPL-Bucky missed by a wide margin (as did the forecast from CU-select) and the allocation was 716 units lower than the allocation derived from the CU-select model.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/thermometer-plot-1} \caption[Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts]{Probabilistic forecasts for new hospital admissions and the inferred resource allocations for COVID-19 on January 3, 2022 for the states with the nine highest hospitalization counts. For each state, the dark black line shows the data observed when the forecast was made and the grey line shows eventually observed counts. The side-by-side shaded regions show the median (solid horizontal line) and 50\%, 80\% and 95\% prediction intervals for the two selected models. The forecasts were made for new hospitalizations on January 3, 2022 (vertical dashed line, with number of hospitalizations indicated by red horizontal line segment at the intersection of the dashed line and the grey line of data). The vertical bars with red and yellow shading show the allocations: the red bar goes from zero to the level of the observed number of hospitalizations while the yellow bar shows how many resources were suggested by the model to be allocated to that location. The allocations (yellow bar) may be more or less than the observed number, therefore exposed red bar indicates unmet need, while the yellow bar extending above the observed number of hospitalizations means that excess resources were suggested for that given location.}\label{fig:thermometer-plot}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/multi-loc-ranks-1} \caption[Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select)]{Component-wise breakdowns of the allocation score (Panel A) and weighted interval score (Panel B), by location for forecasts of hospitalization admissions on January 3, 2022, for two selected models (JHUAPL-Bucky and CU-select). Panel A shows the observed resource need, in this case the observed number of hospitalizations, for each state, along with the hypothetical number of resources allocated to the given location based on the forecasts from each model. The number of available resources was fixed at 15,000 and forecasts from each model were used to determine an optimal allocation strategy before the resource need was known. For most locations the resource need exceeded the resources allocated, indicated by the `observed resource need' bar being larger than the `allocation' bar. Panel B shows the breakdown of the weighted interval score (WIS) into components of underprediction, overprediction and dispersion. Larger values of WIS indicate more error, and the full WIS score for each location can be decomposed into the three components shown here.}\label{fig:multi-loc-ranks}
\end{figure}

\end{knitrout}

\subsubsection{Forecast scores showed differences over time}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/metrics-over-time-1} \caption[Hospital admissions and evaluation metrics over time]{Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the assumed resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The x-axis corresponds to the date for which the prediction was made. AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.}\label{fig:metrics-over-time}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/metrics-correlation-1} \caption[Association of standardized ranks for MWIS and allocation score by model and week]{Association of standardized ranks for MWIS and allocation score by model and week. Each facet of the plot corresponds to one model. Within each facet, each point corresponds to a week. The x- and y-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week. Points corresponding to earlier dates have darker shading. The size of the point corresponds to the observed value on the date for which the prediction was made. Models show different degrees of association between the two metrics.}\label{fig:metrics-correlation}
\end{figure}

\end{knitrout}

@@

AS was computed for a range of $K$ from 200 to 60,000 for forecasts made on December 20, 2020 predicting levels of hospitalizations on January 3, 2021 (Figure \ref{fig:multi-k}A).
These calculations highlight that AS was highest at the $K$ nearest to the observed value of 19,581.
Rankings of AS from all models were fairly stable across a range of values for K, with some crossings, especially at $K$ values further away from the observed value.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/multi-k-1} \caption[Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20]{Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20. Panel A shows, for each model, the AS for values of K between 200 and 60,000 at increments of 200. The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations. Panel B shows two possible weighting functions for the Integrated Allocation Score (IAS). The first (dark blue circles) computes weights proportional to a normal distribution centered at 15,000 (solid vertical gray line) with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000. The second (light orange triangles) uses a uniform weight for all possible values of K. Note that the AS used in earlier sections of the application uses the single fixed value of K=15,000.  Panel C shows how either method of computing the IAS (y-axis) is correlated with AS at K=15,000 (x-axis). Every point represents the AS and IAS for one model. The IAS centered at 15,000 are more closely correlated with the AS values.}\label{fig:multi-k}
\end{figure}

\end{knitrout}

### 2024-07-26 notes/methods_new.rmd
more cleanup

Unit-level decision problem: allocate to potential case in location $a$ or $b$. With a decision variable $d$ we write this as $d=a$ or $d=b$.

- elicit forecast $P(Y_a=1) = p$ where $Y_a=1$ if a case in addition to what is currently provided for at horizon $h$ occurs during the day at horizon $h$ in $a$, and 0 otherwise.  $p$ can in particular come from a distributional incidence forecast $F_a$ via $p = 1-F_a(x)$ where $x$ is the number of cases currently provided for in $h$ days.

@@

We define a loss function for the decision problem that encodes the dilemma faced when balancing the risk of a case of unmet need in $a$ incurred when $d=b$ is chosen against the guaranteed "average" unmet need of $\beta$ in $b$ that is incurred when $d=a$ is chosen. This is the function of the decision $d$ and outcome $Y_a$ 
$$
l(d,Y) = \mathbb{1}\{d=a\}\beta + \mathbb{1}\{d=b\}Y_a,
$$
which picks one of the 4 entries in the table
$$
\begin{array}{rcc}
\text{New case in $a$:}& \text{Yes} & \text{No} \\
\hline
\begin{array}{cr}
\text{allocate to:} & 
\begin{array}{cc}
a \\
b
\end{array}
\end{array}
&
\begin{array}{cc}
\beta \\
1
\end{array}
&
\begin{array}{cc}
\beta \\
0
\end{array}
\end{array}
$$

>>>>

### The unit-level decision problem:

Assume that there are currently resources allocated to care for $x_a$ and $x_b$ many cases in locations $a$ and $b$.  Another resource unit becomes availible and we must decide whether to allocate it to a new potential case in location $a$ or $b$. In terms of a decision variable $d$ the decision is whether to set $d=a$ or $d=b$.

To support our decision making process we elicit forecasts $p_a$ of $P(Y_a=1)$ and $p_b$ of $P(Y_b=1)$ where $Y_a, Y_b = 1$ if a case in addition to what is currently provided for at horizon $h$ occurs during the day at horizon $h$ in $a,b$, and 0 otherwise.

$p_{a,b}$ can in particular come from a distributional incidence forecast $F_a$ via $p = 1-F_a(x)$ where $x$ is the number of cases currently provided for in $h$ days.

@@

We define a loss function for the decision problem that encodes the dilemma faced when balancing the risk of a case of unmet need in $a$ incurred when $d=b$ is chosen against the guaranteed "average" unmet need of $\beta$ in $b$ that is incurred when $d=a$ is chosen. This is the function of the decision $d$ and outcome $Y_a$

$$
l(d,Y) = \mathbb{1}\{d=a\}\beta + \mathbb{1}\{d=b\}Y_a,
$$
which picks one of the 4 entries in the table
$$
\begin{array}{rcc}
\text{New case in $a$:}& \text{Yes} & \text{No} \\
\hline
\begin{array}{cr}
\text{allocate to:} & 
\begin{array}{cc}
a \\
b
\end{array}
\end{array}
&
\begin{array}{cc}
\beta \\
1
\end{array}
&
\begin{array}{cc}
\beta \\
0
\end{array}
\end{array}
$$
